analysis

### 2020-06-10 analysis.Rmd
update

@@

@@

    out <- bind_rows(out, temp)
  } 
  out <- out %>%
    gather(metric, value, -c(sp, id, iteration)) %>% # get mean value across iterations for each metric
    group_by(sp, id, metric) %>%
    summarize(value = mean(value)) %>%
    separate(id, sep = "_", into = c("site", "phase", "year")) %>%
    mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
    left_join(site_data) %>%
    left_join(bb_traits, by = c("sp" = "bb.sp"))
}
```

@@

fl_traits <- read.csv("./processed_data/floral_k_type.csv") %>%
  select(plant.sp, k.type)

@@

### FL traits by elevation -- do we see a sharp change at the tree line?
```{r}
survey_trait_elev <- survey %>%
  group_by(site, year, elev.mean,  k.type) %>%
  summarize(abund = sum(flower.cover)) %>%
  group_by(site, elev.mean, year) %>%
  mutate(div = length(unique(k.type)))

net_trait_elev <- net %>%
  group_by(site, year, elev.mean,  k.type) %>%
  summarize(abund = n()) %>%
  group_by(site, elev.mean, year) %>%
  mutate(div = length(unique(k.type)))

@@

ggplot(survey_trait_elev, aes(elev.mean, log(abund))) +
  geom_point() +
  stat_smooth() +
  facet_wrap(~k.type, scales = "free_y")

ggplot(net_trait_elev, aes(elev.mean, log(abund))) +
  geom_point() +
  stat_smooth() +
  facet_wrap(~k.type, scales = "free_y")

ggplot(survey_trait_elev, aes(elev.mean, div)) +
  geom_point() +
  stat_smooth()

ggplot(net_trait_elev, aes(elev.mean, div)) +
  geom_point() +
  stat_smooth()

ggplot(survey_genus_elev, aes(elev.mean, log(abund))) +
  geom_point() +
  stat_smooth() +
  facet_wrap(~plant.genus, scales = "free_y")

@@

@@

@@

@@

### BB richness ~ elevation
ggplot(abund_div, aes(elev.mean, bb.rich)) +
  geom_point() +
  geom_smooth() +
  facet_grid(year ~ phase)

@@

@@

@@

##### Floral richness -- since year does not seem to change the overall pattern (though perhaps the magnitude), I could use a simpler model, perhaps with year as a random intercept effect
```{r eval=FALSE, include=FALSE}
fl_rich_yday <- gam(fl.rich ~ year +
                    te(yday, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
                    data = filter(abund_div),
                    method = "REML")

@@

plot(fl_rich_yday.vis, allTerms = TRUE) +
  ggtitle("FL richness (yday)") +
  xlab("Day of year") +
  ylab("Elevation")

fl_rich_gdd <- gam(fl.rich ~ year +
                    te(gdd.cum, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
                    data = filter(abund_div),
                    method = "REML")

@@

plot(fl_rich_gdd.vis, allTerms = TRUE) +
  ggtitle("FL richness (gdd)") +
  xlab("Day of year") +
  ylab("Elevation")
```

@@

@@

>>>>

### Rarefied network analysis by floral k.type instead of floral species
```{r}
# Rare_net function: rarefies networks by iteratively subsampling more frequently sampled networks to match the transect count of the least frequently sampled one
# species level
rare_net_sp_k.type.s <- function(net, n, iter) {

@@

  for(i in 1:iter) {
    temp <- net %>%
      select(site, date) %>%
      unique() %>%
      group_by(site) %>%
      sample_n(n) %>%
      left_join(net) %>%
      group_by(site, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
      na.omit() %>%
      group_by(site) %>%
      filter(length(unique(bb.sp)) > 1 & length(unique(k.type.s)) > 1) %>%
      ungroup() %>%
      filter(freq > 1) %>% # remove singletons
      dplyr::select(higher = bb.sp, lower = k.type.s, webID = site, freq) %>% # rename columns to match bipartite's expectations
      data.frame() %>% # convert to data frame
      frame2webs() %>% # convert to to bipartite web
      map(specieslevel, level = "higher", index = c("d", "species strength", "proportional generality")) %>%
      map(rownames_to_column) %>%
      map(as_tibble) %>%
      bind_rows(.id = "site") %>% # Collapse list into big data frame
      rename(sp = rowname) %>%
      mutate(iteration = rep(i, n()))

# network level
rare_net_nt_k.type.s <- function(net, n, iter) {

@@

  for(i in 1:iter) {
    temp <- net %>%
      select(site, date) %>%
      unique() %>%
      group_by(site) %>%
      sample_n(n) %>%
      left_join(net) %>%
      group_by(site, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
      na.omit() %>%
      group_by(site) %>%
      filter(length(unique(bb.sp)) > 1 & length(unique(k.type.s)) > 1) %>%
      ungroup() %>%
      dplyr::select(higher = bb.sp, lower = k.type.s, webID = site, freq) %>% # rename columns to match bipartite's expectations
      data.frame() %>% # convert to data frame
      frame2webs() %>% # convert to to bipartite web
      map(networklevel, weighted = TRUE, level = "higher", index = c("H2", "weighted NODF", "niche overlap", "C score", "generality", "web asymmetry", "weighted connectance")) %>% # We probably care mainly about the BB level of the network
      data.frame() %>%
      t() %>%
      data.frame() %>%
      rownames_to_column() %>%
      dplyr::select(site = rowname, everything()) %>%
      as_tibble() %>%
      mutate(iteration = rep(i, n()))

    out <- bind_rows(out, temp)
  } 
  out <- out  %>%
      gather(metric, value, -c(site, iteration)) %>% # get mean value across iterations for each metric
      group_by(site, metric) %>%
      summarize(value = mean(value)) %>%
      left_join(site_data)
}
```

@@

fl_traits <- read.csv("./processed_data/floral_k_type_nom_fix.csv") %>%
  select(plant.sp, k.type) %>%
  mutate(k.type.s = str_extract(k.type, "[-+]?[0-9]*\\.?[0-9]+"),
         k.type.ss = str_extract(k.type, "[-+]?[0-9]*"))

@@

### FL traits by elevation -- do we see a sharp change at the tree line?
```{r}
survey_trait_elev <- survey %>%
  group_by(site, year, elev.mean,  k.type.s) %>%
  summarize(abund = sum(flower.cover))

survey_trait_elev2 <- survey %>%
  group_by(site, year, elev.mean,  k.type.ss) %>%
  summarize(abund = sum(flower.cover))

net_trait_elev <- net %>%
  group_by(site, year, elev.mean,  k.type.s) %>%
  summarize(abund = n())

net_trait_elev2 <- net %>%
  group_by(site, year, elev.mean,  k.type.ss) %>%
  summarize(abund = n())

ggplot(survey_trait_elev, aes(elev.mean, sqrt(abund))) +
  geom_point() +
  geom_smooth() +
  geom_vline(xintercept = 1750, color = "red") +
  facet_wrap(~k.type.s, scales = "free_y")

ggplot(survey_trait_elev2, aes(elev.mean, sqrt(abund))) +
  geom_point() +
  geom_smooth() +
  geom_vline(xintercept = 1750, color = "red") +
  facet_wrap(~k.type.ss, scales = "free_y")

ggplot(net_trait_elev, aes(elev.mean, sqrt(abund))) +
  geom_point() +
  geom_smooth() +
  geom_vline(xintercept = 1750, color = "red") +
  facet_wrap(~k.type.s, scales = "free_y")

ggplot(net_trait_elev2, aes(elev.mean, sqrt(abund))) +
  geom_point() +
  geom_smooth() +
  geom_vline(xintercept = 1750, color = "red") +
  facet_wrap(~k.type.ss, scales = "free_y")

net_trait_genus_elev <- net %>%
  group_by(site, year, elev.mean, plant.genus, k.type.s) %>%
  summarize(abund = n())

survey_trait_genus_elev <- survey %>%
  group_by(site, year, elev.mean, plant.genus, k.type.s) %>%
  summarize(abund = sum(flower.cover))

ggplot(filter(net_trait_genus_elev, k.type.s == 5.1), 
       aes(elev.mean, sqrt(abund))) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~plant.genus)

ggplot(filter(survey_trait_genus_elev, k.type.s %in% c(5.1)), 
       aes(elev.mean, sqrt(abund))) +
  geom_point() +
  geom_smooth() +
  geom_vline(xintercept = 1750, color = "red") +
  facet_wrap(~plant.genus, ncol = 5) +
  ylab("sqrt(Abundance)") +
  xlab("Elevation")

ggplot(filter(net_trait_genus_elev, k.type.s %in% c(5.1)), 
       aes(elev.mean, sqrt(abund))) +
  geom_point() +
  geom_smooth() +
  geom_vline(xintercept = 1750, color = "red") +
  facet_wrap(~plant.genus, ncol = 5) +
  ylab("sqrt(Visitation frequency)") +
  xlab("Elevation")
```

### Genera and families by elevation
```{r}

@@

#### Plot
ggplot(survey_genus_elev, aes(elev.mean, log(abund))) +
  geom_point() +
  stat_smooth() +
  facet_wrap(~plant.genus, scales = "free_y")

@@

ggplot(abund_div, aes(log(fl.abund), bb.abund, color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  ylab("BB abundance") +
  xlab("log(floral cover)") +
  theme_bw()

@@

ggplot(abund_div, aes(elev.mean, log(fl.per.bb))) +
  geom_point(alpha = 0.5) +
  #stat_summary(aes(y = log(fl.per.bb), group=1), fun.y=mean, colour="black", geom="point",group=1) +
  geom_smooth() +
  facet_grid(.~year, scales = "free") +
  ylab("log(fl-per-bb)") +
  xlab("Elevation") +
  theme_light()

@@

ggplot(filter(abund_div_mean, metric == "fl.per.bb"), aes(elev.mean, log(value), color = year)) +
  geom_point(alpha = 1) +
  #geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
  geom_smooth() +
  ylab("FL per BB") +
  xlab("Elevation") +
  theme_light()

ggplot(filter(abund_div_mean, metric == "fl.per.bb"), aes(elev.mean, log(value))) +
  geom_point(alpha = 1, aes(color = year)) +
  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
  #geom_smooth() +
  ylab("FL per BB") +
  xlab("Elevation") +
  theme_light(16)

@@

### BB richness ~ elevation
ggplot(abund_div, aes(elev.mean, bb.rich)) +
  geom_point() +
  geom_smooth() +
  facet_grid(.~year)

@@

ggplot(abund_div, aes(elev.mean, bb.abund)) +
  geom_point() +
  geom_smooth() +
  facet_grid(.~year, scales = "free")

ggplot(abund_div, aes(elev.mean, bb.abund, color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  xlab("Elevation") +
  ylab("BB abundance")  +
  theme_light(16)

ggplot(abund_div, aes(elev.mean, sqrt(bb.abund), color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  xlab("Elevation") +
  ylab("sqrt(BB abundance)")  +
  theme_light(16)

@@

ggplot(abund_div, aes(elev.mean, fl.abund, color = year)) +
  geom_point(alpha = 0.5) +
  theme_light(16) +
  xlab("Elevation") +
  ylab("Floral abundance")

ggplot(abund_div, aes(elev.mean, log(fl.abund), color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  theme_light(16) +
  xlab("Elevation") +
  ylab("log(Floral abundance)")

ggplot(abund_div, aes(elev.mean, sqrt(bb.abund), color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  xlab("Elevation") +
  ylab("sqrt(BB abundance)")  +
  theme_light(16)

@@

##### Floral richness -- since year does not seem to change the overall pattern (though perhaps the magnitude), I could use a simpler model, perhaps with year as a random intercept effect; some explanat
```{r eval=FALSE, include=FALSE}
### By year
fl_rich_yday <- gam(fl.rich ~ year +
                    te(yday, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
                    data = filter(abund_div),
                    method = "REML")

@@

plot(fl_rich_yday.vis, allTerms = TRUE) +
  ggtitle("FL abundance (gdd)") +
  xlab("Day of year") +
  ylab("Elevation")

### With year as parametric term
fl_rich_yday2 <- gam(fl.rich ~ year +
                    te(yday, elev.mean, bs= "cc", k= 10, m=1),
                    data = filter(abund_div),
                    method = "REML")

fl_rich_yday2.vis <- getViz(fl_rich_yday2)
check.gamViz(fl_rich_yday2.vis)
summary(fl_rich_yday2.vis)

plot(sm(fl_rich_yday2.vis, 1)) + 
  l_fitRaster() + 
  l_fitContour() + 
  l_points(size = 1) +
  ggtitle("FL richness (yday)") +
  xlab("Day of year") +
  ylab("Elevation")

plot(pterm(fl_rich_yday2.vis, 1))

### By year
fl_rich_gdd <- gam(fl.rich ~ year +
                    te(gdd.cum, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
                    data = filter(abund_div),
                    method = "REML")

@@

plot(fl_rich_gdd.vis, allTerms = TRUE) +
  ggtitle("FL richness (gdd)") +
  xlab("GDD") +
  ylab("Elevation")

### With year as parametric term
fl_rich_gdd2 <- gam(fl.rich ~ year +
                    te(gdd.cum, elev.mean, bs= "cc", k= 10, m=1),
                    data = filter(abund_div),
                    method = "REML")

fl_rich_gdd2.vis <- getViz(fl_rich_gdd2)
check.gamViz(fl_rich_gdd2.vis)
summary(fl_rich_gdd2.vis)

plot(sm(fl_rich_gdd2.vis, 1)) + 
  l_fitRaster() + 
  l_fitContour() + 
  l_points(size = 1) +
  ggtitle("FL richness (gdd)") +
  xlab("Day of year") +
  ylab("Elevation")

plot(pterm(fl_rich_gdd2.vis, 1))
```

@@

### Rarefied network analysis by k.type instead of by plant sp
```{r, include = FALSE}
### Create yearly subsets to feed to rare_net
net_2012 <- filter(net, year == "2012")
net_2011 <- filter(net, year == "2011")
net_2010 <- filter(net, year == "2010")

### Species level
spmet_2012 <- rare_net_sp_k.type(net_2012, 9, 20) %>%
  mutate(year = factor(rep("2012", n())))
spmet_2011 <- rare_net_sp(net_2011, 4, 20) %>%
  mutate(year = factor(rep("2011", n())))  
spmet_2010 <- rare_net_sp(net_2010, 3, 20) %>%
  mutate(year = factor(rep("2010", n())))

### Network- and group-level
netmet_rare_2012 <- rare_net_nt_k.type(net_2012, n = 9, iter = 20) %>%
  mutate(year = factor(rep("2012", n())))
netmet_rare_2011 <- rare_net_nt_k.type(net_2011, n = 4, iter = 20)  %>%
  mutate(year = factor(rep("2011", n())))
netmet_rare_2010 <- rare_net_nt_k.type(net_2010, n = 3, iter = 20) %>%
  mutate(year = factor(rep("2010", n())))

### Do BB species differ in their k.type associations?
```{r}
web_by_site <- net %>%
  group_by(site, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
  na.omit() %>%
  group_by(site) %>%
  filter(length(unique(bb.sp)) > 1 & length(unique(k.type.s)) > 1) %>%
  ungroup() %>%
  dplyr::select(higher = bb.sp, lower = k.type.s, webID = site, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs()

visweb(web[[1]])
plotweb(web[[1]])

web <- net %>%
  group_by(year, elev.class2, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
  na.omit() %>%
  unite(webID, c(year, elev.class2), sep = "_") %>%
  dplyr::select(higher = bb.sp, lower = k.type.s, webID, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs()

# 2010 high
visweb(web[[1]])
plotweb(web[[1]])

# 2010 low
visweb(web[[2]])
plotweb(web[[2]])

# 2010 med
visweb(web[[3]])
plotweb(web[[3]])

# 2011 high
visweb(web[[4]])
plotweb(web[[4]])

# 2011 low
visweb(web[[5]])
plotweb(web[[5]])

# 2011 med
visweb(web[[6]])
plotweb(web[[6]])

# 2012 high
visweb(web[[7]])
plotweb(web[[7]])

# 2012 low
visweb(web[[8]])
plotweb(web[[8]])

# 2012 med
visweb(web[[9]])
plotweb(web[[9]])

### Does tongue length class predict k.type association? Remember that B. wurflenii is primarily a nectar robber.
```{r}
library(infer)

kugler_desc <- read_delim("./kugler.tsv", delim = "\t") %>%
  select(1,2)

ggplot(bb_traits, aes(pbl.w)) +
  geom_histogram()

net_cont <- net %>%
  group_by(year, elev.class2, pbl.w.class, k.type.ss) %>%
  summarize(freq = n()) %>%
  group_by(elev.class2, pbl.w.class) %>%
  mutate(prop = freq/sum(freq))

ggplot(net_cont, aes(pbl.w.class, freq, fill = k.type.ss)) +
  geom_col() +
  facet_grid(.~elev.class2)

ggplot(net_cont, aes(pbl.w.class, prop, fill = k.type.ss)) +
  geom_col() +
  facet_grid(.~elev.class2)

chisq_mod <- chisq.test(net$pbl.w.class, net$k.type.ss)
chisq_mod

  c <- net %>%
  select(k.type.s, pbl.w.class) %>%
  mutate(k.type.s = factor(k.type.s),
         pbl.w.class = factor(pbl.w.class)) %>%
  chisq_test(k.type.s ~ pbl.w.class)

@@

```{r}

    out <- bind_rows(out, temp)
  } 
  out <- out %>%
    gather(metric, value, -c(sp, id, iteration)) %>% # get mean value across iterations for each metric
    group_by(sp, id, metric) %>%
    summarize(value = mean(value)) %>%
    separate(id, sep = "_", into = c("site", "phase", "year")) %>%
    mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
    left_join(site_data) %>%
    left_join(bb_traits, by = c("sp" = "bb.sp"))
}
